---
title: "ST558 Project 2"
author: "Shyam Gadhwala & Kamlesh Pandey"
output: 
  github_document:
    toc : True
    
---

This vignette is based on the NASA API. The primary purpose of this vignette is to download data from the API and explore visualization package ggplot for exploratory data analysis purpose. 

[NASA API](https://api.nasa.gov/index.html)

# Requirement

For this project following packages are used.

[httr](https://httr.r-lib.org/) to provide a wrapper function and customized to the demand of modern web APIs.

[jsonline](https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html) to provide flexibility in mapping json and R data

[lubridate](https://lubridate.tidyverse.org/) to manipulate date and time

[ggplot](https://ggplot2.tidyverse.org/) used for creating graphics

[tidyverse](https://www.tidyverse.org/) for data analysis purpose

```{r, warning=FALSE, message=FALSE}
library(httr)
library(tidyverse)
library(jsonlite)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(jpeg)
library(lubridate)
library(GGally)
```

# Asteroid - NeoWs API

NeoWs (Near Earth object Web Service) is a REST API for near earth Asteroid information. Near Earth Objects (NEOs) are comets and asteroids and due to the gravitational attraction of nearby planet they enter into earth gravitational orbit

This API has near earth objects (NEO) tracking and the data set has 99% asteroids and only 1% comet data. 

Parameter information:

Magnitude (H): Asteroid absolute magnitude (in general, smaller H implies larger asteroid diameter). This parameter is undefined for comets.

Close-Approach (CA) Date : Date (TBD) of closest Earth approach.

V Relative : Object velocity relative to the earth

Maximum_Diameter (miles): Estimated maximum diameter in miles

Minimum_Diameter (miles): Estimated minimum diameter in miles


```{r irs, warning=FALSE, message=FALSE}

asteroidData <- function(start_date, end_date, ...){
  baseURL <- 'https://api.nasa.gov/neo/rest/v1/'
  apiKey <- 'igUogzKaubKUi5TTgsbYcdVgU8pICrvizcCrCtY5'
  endpoint <- 'feed'
  
  # URL for GET request
  url <- paste0(baseURL, endpoint, '?startDate=', start_date, '&endDate=', end_date,'&api_key=', apiKey)
  
  res <- GET(url)
  asteroidData <- fromJSON(rawToChar(res$content))
  
  # desired date format to query from the API
  dateFormat <- "%Y-%m-%d"
  
  # check if if the start and end date are in YYYY-MM-DD format
  checkStart <- tryCatch(!is.na(as.Date(start_date, dateFormat)), 
             error = function(err) 
             {TRUE})
  
  checkEnd <- tryCatch(!is.na(as.Date(end_date, dateFormat)), 
             error = function(err) 
             {TRUE})
  
  # check for the correct date format
  if (checkStart == FALSE | checkEnd == FALSE){
    message <- paste('[ERROR..!!] Either your start or end date is not is correct YYYY-MM-DD format')
    stop(message)
  }
  # if both date format start and end are in correct format
  else if(checkStart == TRUE & checkEnd == TRUE){
    
    #if correct format, check if end_date > start_date
    if (ymd(end_date) < ymd(start_date)){
      message <- paste('End date' , end_date, 'should be greater than the start date', start_date)
      stop(message())
      
    } 
      # if end date is greater than the start date
      else if (ymd(end_date) > ymd(start_date)){
        # calculate the difference between the start and end date in days
        diffDate <- difftime(ymd(end_date), ymd(start_date), units = 'days')
        
        if (diffDate > 8){
          message <- paste('[WARNING..!!] The difference between the date range should be less than 8 days', 
                           'for this current date range the API will return future Approach date ')
        }
        
        # API Parameters      
        magnitude <- c()
        dMin  <- c()
        dMax <- c()
        velocity <- c()
        Date <- c()
        missDist <- c()
        orbitBody <- c()
        isHazard <- c()
          
        for (i in 1:length(asteroidData$near_earth_objects)) {
          absMag <- asteroidData$near_earth_objects[[i]]$absolute_magnitude_h
          dMin_  <- asteroidData$near_earth_objects[[i]]$estimated_diameter$miles$estimated_diameter_min
          dMax_ <- asteroidData$near_earth_objects[[i]]$estimated_diameter$miles$estimated_diameter_max
          isHazard_ <- asteroidData$near_earth_objects[[i]]$is_potentially_hazardous_asteroid
          
          tempDate <- c()
          tempVel <- c()
          tempDist <- c()
          tempOrbit <- c()
          
          # for other nested dictionary
          for (j in 1:length(asteroidData$near_earth_objects[[i]]$close_approach_data)){
            # get the relative velocity
            vel <- asteroidData$near_earth_objects[[i]]$close_approach_data[[j]]$relative_velocity$kilometers_per_hour
            # store the velocity in temporary variable
            tempVel <- append(tempVel, vel)
            # get the approach date from the API
            date <- asteroidData$near_earth_objects[[i]]$close_approach_data[[j]]$close_approach_date
            # store approach date in a temporary variable
            tempDate <- append(tempDate, date)
            # get the distance in AU
            dist <- asteroidData$near_earth_objects[[i]]$close_approach_data[[j]]$miss_distance$astronomical
            # store the distance in temporary variable
            tempDist <- append(tempDist, dist)
            # get the orbiting body
            orbit <- asteroidData$near_earth_objects[[i]]$close_approach_data[[j]]$orbiting_body
            # store the orbiting data in temporary variable
            tempOrbit <- append(tempOrbit, orbit)
            
          }
          
          # Append the variables into the corresponding main variable
          missDist <- append(missDist, as.numeric(tempDist))
          Date <- append(Date, tempDate)
          velocity <-append(velocity, as.numeric(tempVel))
          magnitude <- append(magnitude, absMag)
          dMin <- append(dMin, dMin_)
          dMax <- append(dMax, dMax_)
          orbitBody <- append(orbitBody, tempOrbit)
          isHazard <- append(isHazard, isHazard_)
  }
  
      }
  }
  # create a data frame with correct feature name
  aesData <- tibble('Magnitude' = magnitude, 
                    'Minimum_Diameter' = dMin, 
                    'Maximum_Diameter' = dMax, 
                    'Relative_Velocity' = velocity,
                    'Approach_Date' = Date, 
                    'Miss_Distance' = missDist,
                    'Orbiting_Body' = orbitBody,
                    'Is_Potentially_Hazardous_Asteroid' = isHazard)
  
return (list(url = url, data = aesData))
  
}

astDf <- apiSelection(api = "Asteroids - NeoWs", '1994-10-10', '1994-12-10')$data
```


```{r, fig.width= 12 , fig.height=6}
# EDA on Asteroid data

plot1 <- ggplot(astDf, aes(x = Approach_Date, y = Miss_Distance))

plot1 + geom_point(aes(color = Is_Potentially_Hazardous_Asteroid, size = Maximum_Diameter), alpha = 0.7) +
  theme(axis.text.x = element_text(angle = 45), axis.text.y = element_blank())+
  theme(axis.text.y = element_blank(), 
        axis.ticks.y = element_blank()
        )

```

```{r, fig.width= 12 , fig.height=6}

# boxplot for min and max diameter

par(mfrow = c(1,2))
plot2 <- ggplot(astDf, aes(x = Is_Potentially_Hazardous_Asteroid,  y = Minimum_Diameter))

plot2 + 
  geom_boxplot() + geom_point(aes(color = Is_Potentially_Hazardous_Asteroid), position = 'jitter') + 
  labs('Box Plot for Minimum Diameter') + 
  xlab('If Asteroid Hazardous ') + 
  ylab('Minimum Diamter')

```

```{r, fig.width= 12 , fig.height=10, message  =FALSE, warning=FALSE}

library(ggplot2)

#storing only numerical feature column for correlation plot
numericalDf <- astDf[, c(1,2,3,4,6)]

# correlation plot
ggpairs(numericalDf, main = 'Scatter Plot', 
        upper = "blank",
        legends = T,
        color = 'Is_Potentially_Hazardous_Asteroid')

```



# Shyam#####################################

```{r cme, message=FALSE, warning=FALSE}

cmeData <- function(startDate, endDate, speed = 0, halfAngle = 0, ...){
  baseUrl <- 'https://api.nasa.gov/DONKI/'
  apiKey <- 'igUogzKaubKUi5TTgsbYcdVgU8pICrvizcCrCtY5'
  
  checkStart <- !is.na(parse_date_time(startDate, orders = "ymd"))
  if(!checkStart){
    errorMessage <- "Please enter the Start Date in the YYYY-mm-dd format and try again."
    stop(errorMessage)
  }
  
  checkEnd <- !is.na(parse_date_time(endDate, orders = "ymd"))
  if(!checkEnd){
    errorMessage <- "Please enter the End Date in the YYYY-mm-dd format and try again."
    stop(errorMessage)
  }
  
  if (as.Date(startDate) > as.Date(endDate)){
    errorMessage <- "The start date cannot be after the end date. Please enter the dates again."
    stop(errorMessage)
  }
  
  if (speed < 0){
    paste0("Warning: ", "The speed cannot be negative. Proceeding with its default value of 0.")
    speed = 0
  }
  
  if (halfAngle < 0){
    paste0("Warning: ", "The half angle cannot be negative. Proceeding with its default value of 0.")
    halfAngle = 0
  }
  
  targetUrl <- paste0(baseUrl, "CMEAnalysis?", "startDate=", startDate, 
                      "&endDate=", endDate, "&speed=", speed,
                      "&halfAngle=", halfAngle, "&api_key=", apiKey)
  
  jsonContent <- fromJSON(rawToChar(GET(targetUrl)$content))
  
  time <- jsonContent$time21_5
  lat <- jsonContent$latitude
  lon <- jsonContent$longitude
  halfAngle_ <- jsonContent$halfAngle
  speed_ <- jsonContent$speed
  type <- jsonContent$type
  
  cmeDataTibble <- tibble(time = time, latitude = lat, longitude = lon, halfAngle = halfAngle_,speed = speed_, type = type)

  
  return(list(url = targetUrl, data = cmeDataTibble))
}

```


# change cmeSampleData to cmeDf
```{r eda_cme, message=FALSE, warning=FALSE}
cmeSampleData <- cmeData("2017-01-01", "2020-01-01")$data

#cmeSampleData <- cmeData("2019-11-10", "2020-11-10")$data

summary(cmeSampleData %>% select(halfAngle, speed, latitude, longitude))
```


```{r, warning=FALSE, message=FALSE, fig.width= 7.5 , fig.height=6}
img <- readJPEG("img\\sun2.jpeg")

cmeSampleData$type <- as.factor(cmeSampleData$type)

show(ggplot(cmeSampleData, aes(x=latitude, y=longitude)) +
    geom_point(aes(color = type, size = speed)) +
    ylim(-150,150) +
    xlim(-90, 90)  +
  labs(title="Plot showing Corona Mass Ejection on Sun's corona",
        x ="Latitude", y = "Longitude"))
```

```{r inital_plot, warning=FALSE, message=FALSE, fig.width= 7.5 , fig.height=6}

img <- readJPEG("img\\sun2.jpeg")

ggplot(cmeSampleData, aes(x=latitude, y=longitude)) +
    background_image(img) +
    geom_point(aes(color = type, size = speed)) +
    ylim(-180,180) +
    xlim(-90, 90)  +
  labs(title="Plot showing Corona Mass Ejection on Sun's corona",
        x ="Latitude", y = "Longitude")
```


```{r categorizeData, warning=FALSE, message=FALSE}
#cmeSampleData$type <- as.factor(cmeSampleData$type)

cmeSampleData$type <- as.factor(cmeSampleData$type)

speedClassfication <- c("Slow paced", "Medium Paced", "Fast Paced", "Hyper Paced")

cmeSampleData <- cmeSampleData %>% 
  mutate(speedC = as.factor(if_else(speed < 500, speedClassfication[1],
                                            if_else(speed < 1000,  speedClassfication[2],
                                                    if_else(speed < 2000, speedClassfication[3], speedClassfication[4])))))


zones <- c("North-East", "North-West", "South-East", "South-West")
cmeSampleData <- cmeSampleData %>%
  mutate(zone = as.factor(if_else(latitude>=0 & longitude>=0, zones[1],
                        if_else(latitude<=0 & longitude>=0,zones[2],
                                if_else(latitude<=0 & longitude<=0, zones[4],
                                        if_else(latitude>=0 & longitude<=0, zones[3], "Error"))))))

angles <- c("low", "medium", "high")
cmeSampleData <- cmeSampleData %>% 
  mutate(halfAngleC = as.factor(if_else(halfAngle <= 25, angles[1],
                                            if_else(halfAngle <= 45,  angles[2],
                                                     angles[3]))))

cmeSampleData
```

```{r summarize_cfs, message=FALSE, warning=FALSE}

cmeSampleData %>%
  group_by(zone, speedC, type) %>%
  summarize(avgSpeed = mean(speed), sdSpeed = sd(speed),
            avgHalfAngle = mean(halfAngle), 
            sdHalfAngle = sd(halfAngle), count = n()) %>% arrange(zone, speedC)

```

```{r cme_con_tbls, message=FALSE, warning=FALSE, fig.width= 7.5 , fig.height=6}

print(table(cmeSampleData$zone, cmeSampleData$speedC, cmeSampleData$type))

img <- readJPEG("img\\sun2.jpeg")

ggplot(cmeSampleData, aes(x=latitude, y=longitude)) +
    background_image(img) +
    geom_point(aes(color = zone, size = speedC, shape = type)) +
    #scale_shape_discrete(name = "Type", labels = c("S", "C")) +
    #scale_color_manual(values = c("C" = "#37a0bf", "S" = "green")) +
    scale_size_discrete(name = "speed", labels = c(speedClassfication[1], speedClassfication[2], speedClassfication[3], speedClassfication[4])) + 
    ylim(-180,180) +
    xlim(-90, 90) +
  
    annotate(geom="text", x=70, y=150, label=paste0("North-East Region\ncount : ", nrow(cmeSampleData %>% filter(zone==zones[1])), "\navgSpeed : ", round(mean((cmeSampleData %>% filter(zone == zones[1]))$speed), 2),
                                                  "\navgHalfAngle : ", round(mean((cmeSampleData %>% filter(zone == zones[1]))$halfAngle), 2)),
              color="White", size=4) + 
  
  annotate(geom="text", x=-70, y=150, label=paste0("North-West Region\ncount : ", nrow(cmeSampleData %>% filter(zone==zones[2])), "\navgSpeed : ", round(mean((cmeSampleData %>% filter(zone == zones[2]))$speed), 2),
                                                  "\navgHalfAngle : ", round(mean((cmeSampleData %>% filter(zone == zones[2]))$halfAngle), 2)),
              color="White", size=4) + 
  
  annotate(geom="text", x=-70, y=-150, label=paste0("South-West Region\ncount : ", nrow(cmeSampleData %>% filter(zone==zones[4])), "\navgSpeed : ", round(mean((cmeSampleData %>% filter(zone == zones[4]))$speed), 2),
                                                  "\navgHalfAngle : ", round(mean((cmeSampleData %>% filter(zone == zones[4]))$halfAngle), 2)),
              color="White", size=4) + 
  
  annotate(geom="text", x=70, y=-150, label=paste0("South-East Region\ncount : ", nrow(cmeSampleData %>% filter(zone==zones[3])), "\navgSpeed : ", round(mean((cmeSampleData %>% filter(zone == zones[3]))$speed), 2),
                                                  "\navgHalfAngle : ", round(mean((cmeSampleData %>% filter(zone == zones[3]))$halfAngle), 2)),
              color="White", size=4) +
  labs(title="Plot showing zone-wise Corona Mass Ejection statistics",
        x ="Latitude", y = "Longitude")

```

```{r cor, message=FALSE, warning=FALSE}

cor <- cor(cmeSampleData$halfAngle, cmeSampleData$speed)
  
ggplot(cmeSampleData, aes(x=halfAngle, y=speed)) +
    geom_point(aes(color = type)) + 
    geom_smooth(method = "lm")

```

```{r barplot}
ggplot(cmeSampleData, aes(x = type)) + 
  geom_bar(aes(fill = halfAngleC), position = "dodge") + 
  scale_fill_discrete(name = "HalfAngle") + 
  facet_grid(. ~ zone)

```

```{r hist}
dates = c()
for (i in 1:nrow(cmeSampleData)){
  dates <- append(dates, strsplit(cmeSampleData$time[i], split="T")[[1]][1])
}
cmeSampleData$date <- dates

months = c()
for (i in 1:nrow(cmeSampleData)){
  months <- append(months, strsplit(cmeSampleData$date[i], split="-")[[1]][2])
}
cmeSampleData$nummonth <- as.numeric(months)
allmonths <- c("Jan","Feb","Mar",
              "Apr","May","Jun",
              "Jul","Aug","Sep",
              "Oct","Nov","Dec")
cmeSampleData$month <- allmonths[cmeSampleData$nummonth]
cmeSampleData$month <- as.factor(cmeSampleData$month)

cmeSampleData$month <- factor(cmeSampleData$month, ordered = TRUE, levels = c("Jan","Feb","Mar",
              "Apr","May","Jun",
              "Jul","Aug","Sep",
              "Oct","Nov","Dec"))

ggplot(cmeSampleData %>% group_by(month) %>% mutate(count = n()), aes(x=month))+
  geom_histogram(aes(fill=type), stat="count")
```


```{r intial_selection, message=FALSE, warning=FALSE}
apiSelection <- function(api, ...){
  
  if (tolower(api) == tolower("Coronal Mass Ejection (CME) Analysis")){
    cmeData(...)
    
  }
  else if(tolower(api) == tolower("Asteroids - NeoWs")){
    asteroidData(...)
    
  }
  else{
    return("This api is not yet supported. Please select from either 'Coronal Mass Ejection (CME) Analysis' or 'Asteroid (AST)'.")
  }
}

```

